Image Forensics using Reflections on Curved Surfaces

### 2014-02-20 paper/main.tex
paper finished, only minor changes left

\chapter{Prior Work}
\section{Similar Geometric Techniques}
There are orensic techniques similar to reflection consistency that take a geometric approach to shadow consistency. Instead of determining if there is a consistent vanishing point, or a plausible consistent camera location (see Section~\ref{sec:planar-reflections}), these techniques determine if there is a consistent location for a light source. One simple technique for determining shadow consistency, akin to the techniques for reflection consistency, involve drawing rays between points on an object and corresponding points on the shadow. These rays should intersect in a plausible location for the light source (or they should be nearly parallel if the light source is the sun) \cite{farid2010image}. This is dependent on the shadows having easily distinguishable features that can be matched up to the object that is casting the shadow.

@@

% explain that camera parameters are needed
If the camera parameters were known (which is outside the scope of this paper), the normal could be estimated anywhere on the visible part of the sphere. Knowing the camera parameters implies that we know exactly where the center of projection is relative to the image plane. If we fit an ellipse to the apparent contour of the sphere (see Section \ref{sec:perspective-distortion} for details on why it is an ellipse), we get an elliptic cone formed by the center of projection and the ellipse \cite[p. 201]{hartley}. There is then a family of spheres that fit into the cone that could possibly have created the apparent contour (with increasing radius as the sphere gets farther away). Since the normals on any sphere in the family map to the image plane the same way, it is sufficient to know what the cone is to find the normals.

@@

\chapter{Future Work}
%TODO counterforensics
The primary flaw in the test for rays going through the center is that there is clearly some offset in the rays, at least for the test set generated by Blender. It is possible that Blender generates the reflections inaccurately, but since the test images were generated using ray tracing (as Blender 2.37+ defaults), it is not clear how or why that would happen. In order to evaluate the validity of the test, more investigation is required into whether or not Blender truly generates images with reflections accurately, and real-life photos with spherical mirrors should be staged as well.

>>>>

\chapter{Prior Work}
% intro
Prior research into geometric consistency in digital images has dealt most notably with shadows and planar reflective surfaces.
\section{Similar Geometric Techniques}
There are forensic techniques similar to reflection consistency that take a geometric approach to shadow consistency. Instead of determining if there is a consistent vanishing point, or a plausible consistent camera location (see Section~\ref{sec:planar-reflections}), these techniques determine if there is a consistent location for a light source. One simple technique for determining shadow consistency, akin to the techniques for reflection consistency, involve drawing rays between points on an object and corresponding points on the shadow. These rays should intersect in a plausible location for the light source (or they should be nearly parallel if the light source is the sun) \cite{farid2010image}. This is dependent on the shadows having easily distinguishable features that can be matched up to the object that is casting the shadow.

@@

% explain that camera parameters are needed
If the camera parameters were known (which is outside the scope of this paper), the normal could be estimated anywhere on the visible part of the sphere. Knowing the camera parameters implies that we know exactly where the center of projection is relative to the image plane. If we fit an ellipse to the apparent contour of the sphere (see Section \ref{sec:perspective-distortion} for details on why it is an ellipse), we get an elliptic cone formed by the center of projection and the ellipse \cite[p. 201]{hartley}. There is then a family of spheres that fit into the cone that could possibly have created the apparent contour (with increasing radius as the sphere gets farther away). Since the normals on any sphere in the family map to the image plane the same way, it is sufficient to know the cone to find the normals.

@@

One strength of both this test and almost any consistency test is that it should not produce any false positives in theory, as long as the tolerance for genuine images is calibrated correctly. It may miss some false negatives, in the case of an object moving towards or away from the sphere, but any drastic offset from the center is clearly a forgery.

\chapter{Future Work}
% counterforensics
The primary flaw in the test for rays going through the center is that there is clearly some offset in the rays, at least for the test set generated by Blender. It is possible that Blender generates the reflections inaccurately, but since the test images were generated using ray tracing (as Blender 2.37+ defaults), it is not clear how or why that would happen. In order to evaluate the validity of the test, more investigation is required into whether or not Blender truly generates images with reflections accurately, and real-life photos with spherical mirrors should be staged as well.
